### 2022-06-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@66d41abcb568ec783472153faab2d2933e7f2bcd 🚀

\hypertarget{tbl:comparison}{}
\begin{longtable}[]{@{}rccccccc@{}}
\caption{\label{tbl:comparison}Values of four performance metrics, and
five network structure metrics, for 500 independent predictions similar
to the ones presented in fig.~\ref{fig:ecovalid}. The values in
\textbf{bold} indicate the best value for each column (including ties).
Because the values have been rounded, values of 1.0 for the ROC-AUC
column indicate an average \(\ge 0.99\).}\tabularnewline
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
\(Q\)\tabularnewline
\midrule
\endfirsthead
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
\(Q\)\tabularnewline
\midrule
\endhead
Decision tree & 0.85 & 0.92 & 0.97 & 0.12 & 0.21 & 0.76 &
0.31\tabularnewline
BRT & \textbf{0.90} & 0.90 & 0.98 & 0.86 & 0.23 & 0.82 &
0.27\tabularnewline
Random Forest & \textbf{0.90} & \textbf{0.96} & \textbf{1.00} & 0.27 &
\textbf{0.20} & \textbf{0.72} & \textbf{0.32}\tabularnewline
k-NN & 0.80 & 0.91 & 0.95 & 0.58 & 0.24 & 1.0 & 0.18\tabularnewline
Ensemble & 0.88 & 0.94 & \textbf{1.00} & \textbf{0.96} & \textbf{0.20} &
0.75 & 0.31\tabularnewline
Data & & & & & 0.18 & 0.66 & 0.34\tabularnewline
\bottomrule
\end{longtable}

@@

We establish that due to the low prevalence of interactions, even poor
classifiers applied to food web data will reach a high accuracy; this is
because the measure is dominated by the accidentally correct predictions
of negatives. On simulated confusion matrices with ranges of imbalance
that are credible for ecological networks, MCC had the most desirable
behavior, and informedness is a linear measure of classifier skill. By
performing simulations with four models and an ensemble, we show that
informedness and ROC-AUC are consistently high on network data, and that
MCC and PR-AUC are more accurate measures of the effective performance
of the classifier. Finally, by measuring the structure of predicted
networks, we highlight an interesting paradox: the models with the best
performance measures are not the models with the closest reconstructed
network structure. We discuss these results in the context of
establishing guidelines for the prediction of ecological interactions.

TODO informedness and accuracy should be easy to beat, make sure the
model is better than them!

@@

First, because we should have more trust in reported interactions than
in reported absences of interactions, we can draw on previous literature
to recommend informedness as a measure to decide on a threshold (Chicco
et al., 2021); this being said, because informedness is insensitive to
bias, the model performance is better evaluated through the use of MCC
fig.~\ref{fig:biasmccinf}. Because \(F_1\) is monotonously sensitive to
classifier bias fig.~\ref{fig:bias} and network connectance
fig.~\ref{fig:connectance}, MCC should be prefered as a measure of model
evaluation.

@@

Thirdly, regardless of network connectance, maximizing informedness
required a training set balance of about 0.5, and maximizing the MCC
required a training set bias of 0.7 and more. This has an important
consequence in ecological networks, for which the pool of positive cases
(interactions) to draw from is typically small: the most parsimonious
measure (\emph{i.e.} the one requiring to discard the least amount of
information to train the model) will give the best validation potential,
and is probably the informedness (maximizing informedness is the
generally accepted default for imbalanced classification; Schisterman et
al., 2005). This last result suggests that the amount of bias \emph{is}
an hyper-parameter that must be fine-tuned, as using the wrong bias can
lead to under-performing models.

One key element for real-life data that can make the prediction exercise
more tractable is that some interactions can safely be assumed to be
impossible; indeed, a lot of networks admit a stochastic block model as
a good approximation (\emph{e.g.} Xie et al., 2017). In ecological
networks, this can be due to spatial constrains (Valdovinos, 2019), or
to the long-standing knowledge that some links are ``forbidden'' due to
traits (Olesen et al., 2011) or abundances (Canard et al., 2014). The
matching rules (Olito \& Fox, 2015; Strona \& Veech, 2017) can be
incorporated in the model either by adding compatibility traits, or by
\emph{only} training the model on pairs of species that are not likely
to be forbidden links. Besides forbidden links, a real-life case that
may arise is multi-interaction or multi-layer networks (Pilosof et al.,
2017). These can be studied using the same general approach outlined
here, either by assuming that pairs of species can interact in more than
one way (wherein one would train a model for each type of interaction,
based on the relevant predictors), or by assuming that pairs of species
can only have one type of interaction (wherein this becomes a
multi-label classification problem).

@@

>>>>

\hypertarget{tbl:comparison}{}
\begin{longtable}[]{@{}rccccccccc@{}}
\caption{\label{tbl:comparison}Values of four performance metrics, and
five network structure metrics, for 500 independent predictions similar
to the ones presented in fig.~\ref{fig:ecovalid}. The values in
\textbf{bold} indicate the best value for each column (including ties).
Because the values have been rounded, values of 1.0 for the ROC-AUC
column indicate an average \(\ge 0.99\).}\tabularnewline
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) & \(A\)
& Jaccard\tabularnewline
\midrule
\endfirsthead
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) & \(A\)
& Jaccard\tabularnewline
\midrule
\endhead
Decision tree & 0.59 & 0.94 & 0.97 & 0.04 & 0.17 & 0.64 & 0.37 &
\textbf{0.42} & 0.1\tabularnewline
BRT & 0.46 & 0.91 & 0.97 & 0.36 & 0.2 & 0.78 & 0.29 & 0.41 &
0.19\tabularnewline
Random Forest & 0.72 & \textbf{0.98} & 0.99 & 0.1 & \textbf{0.16} &
\textbf{0.61} & 0.38 & \textbf{0.42} & \textbf{0.06}\tabularnewline
k-NN & 0.71 & \textbf{0.98} & 0.99 & 0.02 & \textbf{0.16} &
\textbf{0.61} & \textbf{0.39} & \textbf{0.42} &
\textbf{0.06}\tabularnewline
\emph{Ensemble} & \textbf{0.74} & \textbf{0.98} & \textbf{1.0} &
\textbf{0.79} & \textbf{0.16} & \textbf{0.61} & 0.38 & \textbf{0.42} &
\textbf{0.06}\tabularnewline
\emph{Data} & & & & & 0.16 & 0.56 & 0.41 & 0.42 & 0.0\tabularnewline
\bottomrule
\end{longtable}

@@

We establish that due to the low prevalence of interactions, even poor
classifiers applied to food web data will reach a high accuracy; this is
because the measure is dominated by the accidentally correct predictions
of negatives. On simulated confusion matrices with ranges of imbalance
that are credible for ecological networks, MCC had the most desirable
behavior, and informedness is a linear measure of classifier skill. By
performing simulations with four models and an ensemble, we show that
informedness and ROC-AUC are consistently high on network data, and that
MCC and PR-AUC are more accurate measures of the effective performance
of the classifier. Finally, by measuring the structure of predicted
networks, we highlight an interesting paradox: the models with the best
performance measures are not necessarilly the models with the closest
reconstructed network structure. We discuss these results in the context
of establishing guidelines for the prediction of ecological
interactions.

@@

First, because we should have more trust in reported interactions than
in reported absences of interactions, we can draw on previous literature
to recommend informedness as a measure to decide on a threshold (Chicco
et al., 2021); this being said, because informedness is insensitive to
bias, the model performance is better evaluated through the use of MCC
fig.~\ref{fig:biasmccinf}. Because \(F_1\) is monotonously sensitive to
classifier bias fig.~\ref{fig:bias} and network connectance
fig.~\ref{fig:connectance}, MCC should be prefered as a measure of model
evaluation and comparison.

@@

Third, for the same reasons that should give more weight to PR-AUC than
ROC-AUC, accuracy alone should not be used as a measure of model
performance, but rather as an expectation of how well the model should
behave given the balance in the set on which predictions are made; this
is because, as derived earlier, the expected accuracy for a no-skill
no-bias classifier is \(\rho^2 + (1-\rho)^2\) (where \(\rho\) is the
class balance), which will very easily by large. This pitfall is notably
illustrated in a recent model (Caron et al., 2022) wherein the authors
claim that using a training set of \(n = 10^4\) with only 100 positive
interactions (representing 0.1\% of the total interactions) reached a
good accuracy. Reporting a good accuracy isn't informative is this
accuracy isn't (i) compared to the baseline expected value under the
class balance, and (ii) interpreted in the context of a measure that
isn't sensitive to true negatives (like MCC).

Finally, network connectance (\emph{i.e.} the empirical class imbalance)
should inform the composition of the training and testing set. In the
approach outlined here, we treat the class imbalance as an
hyper-parameter, but \emph{test} the model on a set that has the same
class imbalance as the actual dataset. This is an important change to
the overall methodology so far, as it ensure that the prediction
environment matches the testing environment, and so the values on the
testing set can be directly compared to the values for the actual
prediction. A striking result fig.~\ref{fig:optimbias} is that
Informedness was almost always maximal at 50/50 balance (regardless of
connectance), whereas MCC required \emph{more} positives to be maximized
when connectance \emph{increases}. This has an important consequence in
ecological networks, for which the pool of positive cases (interactions)
to draw from is typically small: the most parsimonious measure
(\emph{i.e.} the one requiring to discard the least amount of
information to train the model) will give the best validation potential,
and is probably the informedness (maximizing informedness is the
generally accepted default for imbalanced classification; Schisterman et
al., 2005). This last result further strengthens the assumption that the
amount of bias \emph{is} an hyper-parameter that must be fine-tuned, as
using the wrong bias can lead to models with excess biass or lower
skill.

One key element for real-life data that can make the prediction exercise
more tractable is that some interactions can safely be assumed to be
impossible; indeed, a lot of networks admit a stochastic block model as
a good approximation (\emph{e.g.} Xie et al., 2017). In ecological
networks, this can be due to spatial constrains (Valdovinos, 2019), or
to the long-standing knowledge that some links are ``forbidden'' due to
traits (Olesen et al., 2011) or abundances (Canard et al., 2014). The
matching rules (Olito \& Fox, 2015; Strona \& Veech, 2017) can be
incorporated in the model either by adding compatibility traits, or by
\emph{only} training the model on pairs of species that are not likely
to be forbidden links; having this information would allow to assemble
training/testing sets that have true negatives, and in this situation,
it may be possible to use the more usual 70/30 split. Besides forbidden
links, a real-life case that may arise is multi-interaction or
multi-layer networks (Pilosof et al., 2017). These can be studied using
the same general approach outlined here, either by assuming that pairs
of species can interact in more than one way (wherein one would train a
model for each type of interaction, based on the relevant predictors),
or by assuming that pairs of species can only have one type of
interaction (wherein this becomes a multi-label classification problem).

@@

\leavevmode\hypertarget{ref-Caron2022AddElt}{}%
Caron, D., Maiorano, L., Thuiller, W., \& Pollock, L. J. (2022).
Addressing the Eltonian shortfall with trait-based interaction models.
\emph{Ecology Letters}, \emph{25}(4), 889--899.
\url{https://doi.org/10.1111/ele.13966}
